---
title: "Data Science HW 3"
author: "Hingling Yu"
date: "2023-10-07"
output: github_document
geometry: margin = 0.85in
fontsize: 12pt
---

## Problem 0

- Set up, with R code hided

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

#### Read dataset

```{r}
data("instacart")

instacart = 
  instacart |> 
  as_tibble()
```

### Answer the questions

- The `instacart` dataset contains `r nrow(instacart)` of rows, and each row representing a single product from an instacart order. Also contains `r ncol(instacart)`columns, which are variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. 

- In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

```{r}
instacart |>
  count(aisle) |> # How many aisles
  arrange(desc(n)) # See what aisles are the most items order from so descending
```

- According to the datas above, there are 134 aisles in total, and fresh vegetables and fresh fruits are the most items ordered from.


#### Make a plot that shows the number of items ordered in each aisle

```{r}
instacart|> 
  count(aisle) |> # count the number of items order in each aisles
  filter(n > 10000) |> # Want aisles with more than 10000 items ordered
  mutate(aisle = fct_reorder(aisle, n)) |> # order aisles by the number of items order
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  labs(
    title = "Number of items ordered in each aisle", # Give a title to the plot
    y = "Number of items(n)") + # Give the y-axis a easier understand label
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) # Twist the angle of text in the x-axis for easier reading
```


#### Table of three most popular items in selected aisles

```{r}
instacart |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |>
  count(product_name) |>
  mutate(rank = min_rank(desc(n))) |># rank by descending so we can easily get the top 3 popular products in each of the aisle we want
  filter(rank < 4) |> # filter the top 3 popular items
  arrange(desc(n)) |> # arrange with the most number of items ordered first
  rename(number_of_times_ordered = n)|> # rename n to number_of_times_ordered
  select(-rank)|> #remove variable rank
  knitr::kable()
```



#### Table of the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) |>
  knitr::kable(digits = 2)

```



## Problem 2

#### Read dataset

```{r}
library(p8105.datasets)
data("brfss_smart2010")

brfss_smart2010 = 
  brfss_smart2010 |>
  tibble()
```


#### Clean data `brfss_smart2010`

```{r}
clean_brfss_smart2010 =
  brfss_smart2010 |>
  janitor::clean_names() |>
  filter(
    topic == "Overall Health", 
    response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) |>
  mutate(
    response = as.factor(response),
    response = fct_relevel(
      response, "Poor", "Fair", "Good", "Very good", "Excellent")) 
  
```


### Answer the questions

#### States were observed at 7 or more locations in 2002 and 2010

```{r}
# states were observed at 7 or more locations in 2002 and 2010
clean_brfss_smart2010 |>
  filter(year %in% c(2002,2010)) |>
  group_by(year, locationabbr)|>
  summarize(num_locations = n_distinct(locationdesc)) |>
  filter(num_locations >= 7)
```
- Based on the datas above, in 2002, Connecticut, Florida, Massachusetts, North Carolina, and Pennsylvania observed more than 7 locations. And in 2010, California, Colorado, Florida, Massachusetts, Maryland, North Carolina, Nebraska, New Jersey, New York, Ohio, Pennsylvania, South Carolina, Texas, and Washington observed more than 7 locations.

#### “spaghetti” plot of average `data_value` over time within a state

```{r}
# Limit to excellent responses, and contains year, state and average data_value cross locations
excellent_brfss = 
  clean_brfss_smart2010 |>
  filter(response == "Excellent") |>
  group_by(year, locationabbr) |>
  summarize(avg_data_value = mean(data_value, na.rm = TRUE))

# Plot of average `data_value` over time within a state
excellent_brfss |>
  ggplot(aes(x = year, y = avg_data_value, group = locationabbr, color = locationabbr)) +
  geom_line() +
  labs(title = "Plot of Average Data Value Over Time",
       x = "Year",
       y = "Average Data Value",
       color = "State") +
  theme(legend.position = "right") +
  viridis::scale_color_viridis(
    discrete = TRUE
  )
```

#### Two-panel plot for the years 2006 and 2010, distribution of `data_value` for responses (“Poor” to “Excellent”) among locations in NY State.


```{r}
brfss_ny_2006_2010 = 
  clean_brfss_smart2010 |>
  filter(locationabbr == "NY",
         year %in% c(2006,2010))

brfss_ny_2006_2010 |>
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  facet_grid(. ~ year) +
  labs(title = "Distribution of Data Value by Response in NY",
       x = "Response",
       y = "Data Value") 
```

## Problem 3

#### Clean the `nhanes_accel` dataset

```{r}
clean_nhanes_accel = 
  read_csv("data/nhanes_accel.csv") |>
  janitor::clean_names() |>
  pivot_longer(                                            
    min1:min1440,
    names_to = "min",
    values_to = "mims",
    names_prefix = "min"
  ) |>
  mutate(
    min = as.numeric(min)
  )
```


#### Clean the `nhanes_covar` dataset
```{r}
clean_nhanes_covar =
  read_csv("data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  filter(age >= 21) |>
  drop_na() |>
  mutate(
    sex = ifelse(sex == 1, "male", "female"),
    education = case_when(
      education == 1 ~ "less than high school",
      education == 2 ~ "high school equivalent",
      education == 3 ~ "more than high school"),
    education = fct_relevel(education, "less than high school", "high school equivalent", "more than high school")
  )

```

#### Combine `nhanes_covar` and `nhanes_accel` datasets

```{r}
merged_nhanes = 
  left_join(clean_nhanes_covar, clean_nhanes_accel, by = "seqn")
```


#### Table for the number of men and women in each education category

```{r}
merged_nhanes|>
  group_by(sex, education) |>
  distinct(seqn) |>
  summarize(count = n()) |>
  knitr::kable()
```

#### Plot of age distribution by gender and education

```{r}
merged_nhanes|>
  ggplot(aes(x = education, y = age, fill = sex)) +
  geom_boxplot() + 
  labs(title = "Age distribution by Gender and Education", 
       y = "Age", 
       x = "Education") 

str(merged_nhanes)
```

- Within the "less than high school" level, females have a slightly higher median age than males. In the "high school equivalent" level, the median age for females is higher compared to males. For those with "more than high school" education, the median ages of males and females are close around the age of 40s.The median age tends to decrease with higher education for both genders. Also, both males and females exhibit similar age distributions across the different education levels.  Across all education levels, the age range of males and females exhibits considerable overlap, but the interquartile range for males appears broader, especially in the "high school equivalent" level.

#### Plot of total activities vs age by sex and education level

```{r, warning=FALSE}
# Aggregate activity by sequence number

total_activity_nhanes =
  merged_nhanes |>
  group_by(seqn, sex, age, education) |>
  summarize(
    total_activity = sum(mims)
    )

total_activity_nhanes |>
  ggplot(aes(x = age, y = total_activity, color = sex)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ education) +
  labs(title = "Total Activity vs Age by Sex and Education Level", 
       x = "Age", 
       y = "Total Activity") +
  theme_minimal()
```

- According to the plot, total activity tends to decrease with age. It is noticeable across all education levels, where younger individuals generally have higher total activity compared to their older counterparts. The decline is particularly sharp after the age of 60, suggesting a significant reduction in physical activity as people approach later stages of life.

- There are noticeable differences between men and women. In the "less than high school" education level, males have higher activity level than females between the age 40 to 80. In both "high school equivalent" and "more than high school" education level, females appear to have higher total activity than males, especially in the age range of 20 to 60. Interestingly, the female's total activity level increase again for a little bit between the age 60 to 70 in the "high school equivalent" education level. Among those with "more than high school" education, the activity levels drop significantly only after the age of 60, and the differences between male and female activity levels seem less consistent, especially in the age range of 40 to 60.

#### Plot of 24-hour activity time courses by education level

```{r}
merged_nhanes |>
  ggplot(aes(x = min, y = mims, group = seqn, color = sex)) +
  geom_line(alpha = .2) +
  geom_smooth(aes(group = sex), se = FALSE) +
  facet_grid(. ~ education) +
  labs(
    title = "24-hour Activity Time Courses by Education Level",
    x = "Time (in mins)",
    y = "Avtivity Times",
  ) 
```

- According to the plot, a peak of activity is observed around 750 mins across all education levels. This suggests participants are most active during midday or early afternoon. Additionally, activity levels are low during the early and late hours, aligning with typical sleeping times.

- For both "less than high school" and "high school equivalent" education levels, males and females exhibit similar activity patterns throughout the day. However, in the "more than high school" category, females are slightly more active than males between morning to noon time, while males surpass females in activity around dinner time.


